A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-11-12 paper/paper.tex
paper: fix inconsistencies (WIP)

In there, the process of construction of the type of $X_1$ from that of $X_0$, and from the type of $X_2$ to that of $X_1$, and so on, is similar to iteratively applying a binary parametricity translation. The parametricity which we consider interprets a type $A$ by a family $A_\kstar$ of edges in a graph over $A$, and a term $t: A$ as an edge in $A_\kstar(t)(t)$. To each type constructor is then associated the construction of a graph. In particular, the type of types $\U$ is interpreted as the family of families ${\U}_\kstar$, which takes $A_L$ and $A_R$ in $\U$ and returns the type $A_L \rightarrow A_R \rightarrow \U$ of families of edges over $A_L$ and $A_R$. Also, for $A$ interpreted by $A_\kstar$ and $B$ interpreted by $B_\kstar$, a dependent function type $\Pi a: A.\, B$ is interpreted as the graph $(\Pi a: A.\, B)_\kstar$ that takes two functions $f_L$ and $f_R$ of type $\Pi a: A.\, B$, and expresses that these functions map related arguments in $A$ to related arguments in $B$:
\begin{align*}
  (\Pi a: A.\, B)_\kstar(f_L)(f_R) \; \defeq \; \Pi a_L: A.\, \Pi a_R: A.\, \Pi a_\kstar: A_\kstar(a_L)(a_R).\, B_\kstar(f_L(a_L))(f_R(a_R))
\end{align*}

In particular, for $X: \U$, applying the parametricity translation is about canonically associating to $X$ an inhabitant $X_\kstar$ of ${\U}_\kstar(X)(X)$ i.e. of $X \rightarrow X \rightarrow \U$. In turn, applying the parametricity translation to $X_\kstar: X \rightarrow X \rightarrow \U$ is about canonically associating to $X_\kstar$ an inhabitant $X_{\kstar\kstar}$ of $(X \rightarrow X \rightarrow \U)_\kstar(X_\kstar)(X_\kstar)$ i.e. of:
\begin{align*}
  \Pi x_{LL}: X.\, \Pi x_{LR}: X.\, X_\kstar(x_{LL})(x_{LR})             \\
  \rightarrow \Pi x_{RL}: X.\, \Pi x_{RR}: X.\, X_\kstar(x_{RL})(x_{RR}) \\ \rightarrow X_\kstar(x_{LL}, x_{RL}) \rightarrow X_\kstar(x_{LR})(x_{RR})  \rightarrow \U
\end{align*}
which hints us at how the sequence $X_0$, $X_1$, $X_2$ can be seen as a sequence of inhabitants of the iteration of binary parametricity applied to an initial $X: \U$:
\begin{equation*}
  \begin{array}{llllr}
    X_0 & \defeq & X                & : & \U                                               \\
    X_1 & \defeq & X_\kstar         & : & {\U}_\kstar(X)(X)                                \\
    X_2 & \defeq & X_{\kstar\kstar} & : & ({\U}_\kstar(X)(X))_{\kstar}(X_\kstar)(X_\kstar) \\
    \ldots
  \end{array}
\end{equation*}

>>>>

In there, the process of construction of the type of $X_1$ from that of $X_0$, and from the type of $X_2$ to that of $X_1$, and so on, is similar to iteratively applying a binary parametricity translation. The parametricity which we consider interprets a type $A$ by a family $A_\kstar$ over $A \times A$, what can be seen as a graph whose vertices are in $A$. To each type constructor is thus associated the construction of a graph. In particular, the type of types $\U$ is interpreted as the family of families ${\U}_\kstar$, which takes $A_L$ and $A_R$ in $\U$ and returns the type $A_L \rightarrow A_R \rightarrow \U$ of families over $A_L$ and $A_R$. Also, for $A$ interpreted by $A_\kstar$ and $B$ interpreted by $B_\kstar$, a dependent function type $\Pi a: A.\, B$ is interpreted as the graph $(\Pi a: A.\, B)_\kstar$ that takes two functions $f_L$ and $f_R$ of type $\Pi a: A.\, B$, and expresses that these functions map related arguments in $A$ to related arguments in $B$:
\begin{align*}
  (\Pi a: A.\, B)_\kstar(f_L)(f_R) \; \defeq \; \Pi a_L: A.\, \Pi a_R: A.\, \Pi a_\kstar: A_\kstar(a_L,a_R).\, B_\kstar(f_L(a_L),f_R(a_R))
\end{align*}

In particular, for $X: \U$, applying the parametricity translation is about canonically associating to $X$ an inhabitant $X_\kstar$ of ${\U}_\kstar(X,X)$ i.e. of $X \times X \rightarrow \U$. In turn, applying the parametricity translation to $X_\kstar: X \times X \rightarrow \U$ is about canonically associating to $X_\kstar$ an inhabitant $X_{\kstar\kstar}$ of $(X \times X \rightarrow \U)_\kstar(X_\kstar,X_\kstar)$ i.e. of:
\begin{align*}
  \Pi x_{LL}: X.\, \Pi x_{LR}: X.\, X_\kstar(x_{LL},x_{LR})              \\
  \rightarrow \Pi x_{RL}: X.\, \Pi x_{RR}: X.\, X_\kstar(x_{RL})(x_{RR}) \\ \rightarrow X_\kstar(x_{LL}, x_{RL}) \times X_\kstar(x_{LR},x_{RR})  \rightarrow \U
\end{align*}
which hints us at how the sequence $X_0$, $X_1$, $X_2$ can be seen as a sequence of inhabitants of the iteration of binary parametricity applied to an initial $X: \U$:
\begin{equation*}
  \begin{array}{llllr}
    X_0 & \defeq & X                & : & \U                                               \\
    X_1 & \defeq & X_\kstar         & : & {\U}_\kstar(X)(X)                                \\
    X_2 & \defeq & X_{\kstar\kstar} & : & ({\U}_\kstar(X)(X))_{\kstar}(X_\kstar)(X_\kstar) \\
    \ldots
  \end{array}
\end{equation*}
